Assessment of forestry road network projection methods using the “roads” package in R

### 2022-10-04 analysis/paper/paper.Rmd
update raw data because some roads were missing on the edge
Also updated R packages and fixed some things
update projection method example figure

---
title: "Assessment of forestry road network projection methods using the “roads” package in R"
author:
  - name: Josie Hughes
    email: josie.hughes@ec.gc.ca
    institute: [ECCC]
    correspondence: true
  - name: Sarah Endicott
    email: sarah.endicott@ec.gc.ca
    institute: [ECCC]
    correspondence: false
institute:
  - ECCC: Environment and Climate Change Canada 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-forest-ecology-and-management.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
editor_options: 
  chunk_output_type: console
---

@@

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300,
  cache = TRUE
)
```

@@

```{r fig1, fig.cap="Revelstoke timber supply area in British Colombia, Canada. The colour shows the cost of building roads across each pixel."}
tsaBoundary <- read_sf(here(data_path_raw, "tsa27_boundaries.gpkg"))
#cost surface raster layer
bc_cost_surface <- terra::rast(here(data_path_raw, "cost_surface_bc_ha.tif"))
lakeValue <- 65000

@@

canada <- canada %>% 
  st_as_sf() %>% 
  filter(NAME_0 == "Canada") %>% 
  st_transform(st_crs(tsaBoundary)) %>% 
  st_simplify(dTolerance = 10000)

@@

```{r fig2, fig.height=5, fig.width=fig_widths["two"], fig.cap="Example of the different projection methods and sampling strategies."}
knitr::include_graphics(here("analysis/figures/projection_methods_figure.jpg"))
```

>>>>

---
title: "Assessment of forestry road network projection methods using the “roads” package in R"
author:
  - name: Josie Hughes
    email: josie.hughes@ec.gc.ca
    institute: [ECCC]
    correspondence: true
  - name: Sarah Endicott
    email: sarah.endicott@ec.gc.ca
    institute: [ECCC]
    correspondence: false
institute:
  - ECCC: Environment and Climate Change Canada 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-forest-ecology-and-management.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
editor_options: 
  chunk_output_type: inline
---

@@

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

@@

```{r fig1, fig.cap="Revelstoke timber supply area in British Colombia, Canada. The colour shows the cost of building roads across each pixel.", cache=TRUE}
tsaBoundary <- read_sf(here(data_path_raw, "tsa27_boundaries.gpkg"))
#cost surface raster layer
bc_cost_surface <- terra::rast(here(data_path_raw, "cost_surface_bc_ha.tif"))
lakeValue <- 65000

@@

canada <- canada %>% 
  st_as_sf() %>% 
  filter(COUNTRY == "Canada") %>% 
  st_transform(st_crs(tsaBoundary)) %>% 
  st_simplify(dTolerance = 10000)

@@

```{r, cache=TRUE}
# Example area for demonstrating different methods

library(raster)
library(dplyr)
library(sf)
library(roads)
library(terra)
library(tmap)
library(purrr)
library(caribouMetrics)
library(pfocal)

pal <- met.brewer(pal_nm, 8) %>% .[c(1,3,5,7)]

# read in example area data
exRoads <- read_sf(here(data_path_drvd, "testing_ex_roads.gpkg"))
tsb <- read_sf(here(data_path_drvd, "testing_tsb.gpkg"))
cutblocks <- read_sf(here(data_path_drvd, "testing_cutblocks.gpkg"))
tsaCost <- raster(here(data_path_drvd, "testing_cost.tif"))
roads <- read_sf(here(data_path_drvd, "testing_obs_roads.gpkg"))

sampleDens <- c(low,high,low,high,low)
sampleType <- c("regular","regular","random","random","centroid")
paramTable <- tibble(sampleType, sampleDens,
                     runTime = vector("list", length(sampleDens)),
                     output = vector("list", length(sampleDens)),
                     roadDisturbance = vector("list", length(sampleDens)),
                     roadDensity = vector("list", length(sampleDens)),
                     roadPresence = vector("list", length(sampleDens)),
                     distanceToRoad = vector("list", length(sampleDens)),
                     forestryDisturbance = vector("list", length(sampleDens))) %>%
  distinct()

# re-run projections for just the small area with different parameters
mstProj <- projectAll(tsbs = tsb, paramTable = paramTable,
                      costSurface = tsaCost,
                      cutblocks = cutblocks,
                      existingRoads = exRoads,
                      fileLocation = paste0(data_path_drvd, "for_fig/mst"),
                      method = "mst")

dlcpProj <- projectAll(tsbs = tsb,
                      paramTable = paramTable %>% filter(sampleType == "regular"),
                      costSurface = tsaCost,
                      cutblocks = cutblocks,
                      existingRoads = exRoads,
                      fileLocation = paste0(data_path_drvd, "for_fig/dlcp"),
                      method = "dlcp")

lcpProj <- projectAll(tsbs = tsb,
                       paramTable = paramTable %>% filter(sampleType == "centroid"),
                       costSurface = tsaCost,
                       cutblocks = cutblocks,
                       existingRoads = exRoads,
                       fileLocation = paste0(data_path_drvd, "for_fig/lcp"),
                       method = "lcp")

allProj <- bind_rows(mstProj, dlcpProj, lcpProj, .id = "method") %>%
  arrange(desc(sampleType), sampleDens)

allMaps <- purrr::map(
  1:nrow(allProj),
  ~ qtm(tsaCost, raster.style = "cont", raster.palette = "Greys",
        raster.alpha = 0.5)+
    qtm(cutblocks, fill = NULL, borders = pal[1], borders.lwd = 1.5)+
    qtm(roads, lines.col = pal[3], lines.lwd = 2)+
    qtm(read_sf(here(allProj$output[[.x]])), lines.col = pal[2], lines.lwd = 2)+
    qtm(exRoads, lines.col = "black", lines.lwd = 2)+
    tm_layout(legend.show = FALSE,
              main.title = paste(ifelse(allProj$method[.x]== 1, "MST",
                                        ifelse(allProj$method[.x]== 2, "DLCP", "LCP")),
                                 allProj$sampleType[.x],
                                 ifelse(allProj$sampleType[.x] == "centroid", "",
                                        allProj$sampleDens[.x])),
              main.title.size = 0.75))

tmap_save(tmap_arrange(allMaps, ncol = 2),
          here("analysis/figures/projection_methods_figure.png"),
          dpi = 300, height = 8, width = 3.543)
```

```{r fig2, fig.cap="Example of the different projection methods and sampling strategies. Target cutblocks are shown by grey lines, existing roads by black lines, projected roads by green lines, observed roads by red lines, and the cost of road building by the grey shading."}
knitr::include_graphics(here("analysis/figures/projection_methods_figure.png"))
```
